---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r}
library(tidyverse)

# Prevent scientific notation and display results with one significant digit
options(scipen = 999, digits=1)
```

## Loading and preprocessing the data

```{r}
data <- read_csv("activity.csv")
```

## What is mean total number of steps taken per day?

```{r}
total_steps_per_day <- data %>%
    group_by(date) %>%
    summarize(steps = sum(steps)) %>%
    na.omit()

ggplot(data = total_steps_per_day, aes(steps)) +
    geom_histogram(binwidth = 1000)
```

The mean number of steps taken per day was:

```{r}
mean(total_steps_per_day$steps)
```

The median number of steps taken per day was: 
```{r}
median(total_steps_per_day$steps)
```

## What is the average daily activity pattern?

```{r}
average_steps_by_interval <- data %>%
    group_by(interval) %>%
    summarize(steps = mean(steps, na.rm = TRUE))

ggplot(data = average_steps_by_interval, aes(x = interval, y = steps)) +
    geom_line()
```

```{r}
max_interval <- average_steps_by_interval %>%
    filter(steps == max(steps))
```

The 5-minute interval with the highest average number of steps was:

```{r}
max_interval$interval
```

## Imputing missing values

The total number of rows with missing values in `data` is:

```{r}
data$steps %>% is.na %>% sum
```

We will impute missing values by substituting the median of each 5-minute interval. Do this using the awesome `simputation` package.

```{r}
library(simputation)

data_imputed <- data %>%
    impute_median(steps ~ interval)

total_steps_per_day_imputed <- data_imputed %>%
    group_by(date) %>%
    summarize(steps = sum(steps))

ggplot(data = total_steps_per_day_imputed, aes(steps)) +
    geom_histogram(binwidth = 1000)
```

After using imputation, the mean number of steps taken per day was: 
```{r}
mean(total_steps_per_day_imputed$steps)
```

The median number of steps taken per day was:
```{r}
median(total_steps_per_day_imputed$steps)
```

Both of these values are lower than in the original data set.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
data_imputed <- data_imputed %>%
    mutate(
        weekend = ifelse(weekdays(date) %in% c("Saturday", "Sunday"), 
                         "weekend", "weekday"),
        weekend = as.factor(weekend)
    )

average_steps_by_interval_weekend <- data_imputed %>%
    group_by(interval, weekend) %>%
    summarize(steps = mean(steps, na.rm = TRUE))

ggplot(data = average_steps_by_interval_weekend, aes(x = interval, y = steps)) +
    geom_line() +
    facet_wrap (~weekend, nrow = 2)

```
